---
title: "Assignment 4"
author: "Michelle Lepori-Bui"
date: "11/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(lubridate)
library(kableExtra)
library(effsize)
library(broom)
library(ggbeeswarm)

hares <-read_csv(here("data","Bonanza Creek snowshoe hare data.csv"))
```

# Title

## Introduction

4 -5 sentences about what the study is exploring including basic background info suject site and study with citations
brief overview of waht is included in the repot

## Data and analyses

3-4 sentences friefly describe the data including citation and credit 
provide and overview of the main analysis, stat methods, and tools

##exploratory findings

1. annual juvenile hare trap counts
count the total number of juvenile hare trappings (age == "j") during each year of the study include all grid sites
creat a finalized data visualization of the counts by eyar with figure caption
```{r}
juveniles <- hares %>% 
  filter(age == "j") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) 
juvenile_counts <- juveniles %>% 
  count(year)

ggplot(data = juvenile_counts, aes(x = year, y = n)) +
  geom_point()
```

3-4 sentences below figure describe major takeawyas from graph about total snowshoe hare trappings at the site

consider:
min/max counts, mean/median annual number of hares trapped
general/ notable trends
fig represetns absolute counts. effort (# of dats and number of traps each year) could impact
what should you do and how could you standardie the j hare population

use:
lubridate()
count()


2. Visualize juvenile hare weights

```{r}
ggplot(data = juveniles, aes(x = sex, y = weight)) +
  geom_beeswarm(aes(color = sex)) +
  facet_wrap(~grid) +
  
  
  geom_point(data = penguin_f_summary, 
             aes(x = species, y = mean_flipper_length),
             size = 4) +
  geom_errorbar(data = penguin_f_summary,
               aes(x= species,
                   ymax = mean_flipper_length + sd_flipper_length,
                   ymin = mean_flipper_length - sd_flipper_length,),
               width = 0.1)
```




3. Juvenile weight comparison 

**Table 1.** Caption: 

```{r}
juveniles_summary <- juveniles %>% 
  filter(!is.na(weight)) %>% 
  group_by(sex) %>% 
  summarize(mean_weight = mean(weight),
            sd_weight = sd(weight),
            sample_size = n())
juveniles_table <- juveniles_summary %>% 
  kable(col.names = c("Sex", "Mean weight (g)", "Standard deviation (g)", "Sample size")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
juveniles_table
```

```{r, include=FALSE}
juveniles_mf <- juveniles %>% 
  filter(!is.na(sex))

# Histograms
ggplot(data = juveniles_mf, aes(x = weight)) +
  geom_histogram(bins = 12) +
  facet_wrap(~sex)

# QQ Plots
ggplot(data= juveniles_mf, aes(sample = weight)) +
  geom_qq() +
  facet_wrap(~sex)
```


```{r}
juveniles_f <- juveniles %>% 
  filter(!is.na(weight)) %>%
  filter(sex == "f") %>% 
  pull(weight)

juveniles_m <- juveniles %>% 
  filter(!is.na(weight)) %>%
  filter(sex == "m") %>% 
  pull(weight)

cohen.d(juveniles_f, juveniles_m)

juveniles_ttest <- t.test(juveniles_f, juveniles_m)
juveniles_ttest

```


2-3 sentence comperhensice statement of the means comparison includeing 
-actual differnce in means
- effect size (cohesn d)
- outcome of2 sample ttest




4. Relationship between juvenile weight and hind foot length

hindft mm vs weight g
ALL juvenile hares
 figure of foot length vs weight
 caption
 
linear regressio
check model diagnostics

describe the outcome
slope interpretation
r2 value
pearsons r correlation

are the assumptions of linear regression a concern?
additonal thoughts about how well or not the model fits
what worries you?




### Summary


### Citations


## END NO CODE OR WARNINGS!





